The non-random assembly of functional motifs in  \nplant-pollinator networks

### 2022-03-29 Manuscript/Main_text/Main_text_Functional_Ecology.Rmd
Edits

We calculated which plant and floral visitor groups were over- or under-represented in the different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. To estimate the position frequencies of each plant and floral visitor group in a given network, we added the frequencies of those species that belong to the same group, and then, we normalised the resulting frequencies by dividing them by the total number of times that a corresponding group appears in any position within the same motif size class (i.e., number of species of a motif). Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did with motif frequencies. To outline the general patterns of position frequencies across networks and groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the group identifier as an explanatory variable and the study identifiers as a random intercept. Hence, we assessed the average motif frequency per species group, after controlling for the variation at the study level. Finally, we visualized with the help of the package _ComplexHeatmap_ version _2.6.2_ [@gu2016] over- and under-representation of plant and floral visitor groups on the different motif positions.

To understand if the different plant and floral visitor groups are associated with motif positions that represent different network roles we characterized the roles of each motif positions for plant and floral visitor groups in terms of their specialization and number of indirect interactions. To that end, we calculated the specificity of each motif positions as a measure of specialization and we used the number of indirect interactions of the different motif positions to evaluate the association of these two position metrics and the group frequencies for the different motif positions. First, specificity ($s_i$) was calculated for each motif position following @poisot2015 as $s_i = L - l_i / L - 1$, where $L$ is the total number of groups and $l_i$ the number of interaction partners of the group $i$ in a given motif. Values of the vector s, range between 1, complete specialization (1 partner within the motif) and 0, complete generalism (all possible partners within the motif). Then, we explored through linear models the association between the frequency of the different groups (response variable) and specificity (predictor) [NO RANDOMS HERE? WHY](). Second, we calculated the number of indirect interactions for each motif position by accounting only for the indirect effect of two species by a third species. That is, we explored exclusively indirect interactions within plants though shared floral visitors or within floral visitors though shared plants.  [NO ANALYSIS HERE? WHY? HOW DO YOU USE THIS VALUE?]()

>>>>

We calculated which plant and floral visitor groups were over- or under-represented in the different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. To estimate the position frequencies of each plant and floral visitor group in a given network, we added the frequencies of those species that belong to the same group, and then, we normalised the resulting frequencies by dividing them by the total number of times that a corresponding group appears in any position within the same motif size class (i.e., 2, 3, 4 or 5 species motifs). Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did with motif frequencies. To outline the general patterns of position frequencies across networks and groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the group identifier as an explanatory variable and the study identifiers as a random intercept. Hence, we assessed the average motif frequency per species group, after controlling for the variation at the study level. Finally, we visualized with the help of the package _ComplexHeatmap_ version _2.6.2_ [@gu2016] over- and under-representation of plant and floral visitor groups on the different motif positions.

To understand if the different plant and floral visitor groups are associated with motif positions that represent different network roles, we characterized the roles of each motif positions for plant and floral visitor groups in terms of their specialization and number of indirect interactions. To that end, we calculated the specificity of each motif position following @poisot2015 as $s_i = L - l_i / L - 1$, where $L$ is the total number of groups and $l_i$ the number of interaction partners of the group $i$ in a given motif. Values of the vector $si$, range between 1, complete specialization (1 partner within the motif) and 0 complete generalism (all possible partners within the motif). Then, we calculated the number of indirect interactions for each motif position and trophic guild (i.e., plants or floral visitors) as $z_i-1$ where $z_i$ is the total number of nodes for a trophic guild in a given motif. Because all nodes within a motif are connected, to obtain the total number of indirect interactions for a given position we only have to substract to $z_i$ the motif position of interest that cannot interact indirectly with itself. Finally, we analysed in a single model how specialization and the number of indirect interactions are associated with the observed percentiles for each trophic guild and motif position (see paragraph above). For this, we implemented a GLMM model with the help of the _glmmTMB_ package [@magnusson2017] version _1.0.2.1_ with a Beta distribution because percentiles have exclusively  positive values. The variation at the study system was controlled with the help of the study identifier as a random factor.
